A GRASP algorithm for the concrete delivery problem.

### 2023-08-08 Article/article.tex
save article

The procedure $SimulateVisit$ takes as parameters a partial solution, a delivery node $d$, a driver $k$, and a numeric value $\lambda$. Its goal is to find the earliest time at which the delivery must start at $d$ and the best position within the route of $k$. When planning the visit of customer $i$ to node $d$ of order $o$ with driver $k$, we first compute his expected delivery time $v_d$ as follows:

@@

    {\setstretch{1}
        %  {\small
        \begin{algorithm}[htbp]
            \caption{Customer-Based insertion algorithm }
            \label{alg:greedyCustIns}
            \DontPrintSemicolon
            \LinesNumbered
            \setcounter{AlgoLine}{0}
            \KwIn{  $S$: empty solution $S$, $K$: vehicles set \hspace{15mm} \\$\Pi$: custom sorted customers list }

@@

            $delay \leftarrow 0$

@@

For each delivery $d$ in $CL'$, the algorithm determines the best available driver $k$ and adds the corresponding loading and unloading task $(L^k_d, U^k_d)$ to a list of tasks. This is done using the $SimulateVisit$ procedure described earlier. Next, the algorithm constructs a $RCL$ from the list of tasks. If it is empty, the algorithm proceeds to the next iteration. Otherwise, it randomly selects a pair of tasks from the RCL, inserts them into $S$, and updates the remaining order and customer quantities. If an order $o$ is not fulfilled after visiting $d^j_{o}$, we add its next delivery node $d^{j+1}_{o}$ to $CL$.  Otherwise, if the customer has remaining orders, we select the first delivery node $d^{0}_{o'}$ of another randomly selected order $o'$ and add it to $CL$.

@@

\subsection{ Local Search}

We use four local search operators to improve the solution obtained after the construction phase.

\subsubsection*{Swap load operator}

The first operator exchanges loads between two deliveries of the same order. This move affects the loading operations and can potentially improve the number of scheduled orders.

\subsubsection*{Swap driver operator}

This operator investigates different driver assignments by swapping the drivers assigned to two delivery nodes in the current solution.

\subsubsection*{Relocate load operator}

This operator moves a delivery load that was originally split between two drivers to a single driver, if that driver's capacity allows.
%   By consolidating the load into a single vehicle, we can potentially optimize the routing and reduce the total travel distance. This operator helps in better utilizing the available capacity of the vehicles and improving the overall efficiency of the solution.

\subsubsection*{Remove and reschedule operator}

This operator takes a customer $i$ not scheduled in the current solution. It then identifies a set $C$ of scheduled customer's with the same time slot as $i$. Elements of $C$ are removed and attempted to be rescheduled along with $i$.

% The application of these local search operators is performed iteratively, allowing for multiple rounds of neighborhood exploration and solution refinement. Each operator is applied sequentially, and the resulting solution is evaluated to determine if an improvement has been achieved. The process continues until no further improvements can be made, indicating that a locally optimal solution has been reached.

@@

After the construction and local search steps, it may happen that some customer orders cannot be scheduled within a solution. The goal of the path relinking procedure is to use each customer's scheduling information from existing solutions and combine them effectively to construct a solution that schedules all customers. This algorithm iterates through each element of a pool of solutions and attempts to insert the customer's deliveries into a newly constructed solution.

\section{ Experimental results}
\label{comp_exp}

>>>>

The $SimulateVisit$ procedure takes a partial solution, a delivery node $d$, a driver $k$ and a numeric value $\lambda$ as parameters. Its goal is to find the earliest time when an unloading service can start at $d$ and the best position within the route of $k$. When planning to visit node $d$ of order $o$ and customer $i$ with driver $k$, we first compute his expected delivery time $v_d$ as follows:

@@

    {\setstretch{1}
        %  {\small
        \begin{algorithm}[]
            \caption{Customer-Based insertion algorithm }
            \label{alg:greedyCustIns}
            \DontPrintSemicolon
            \LinesNumbered
            \setcounter{AlgoLine}{0}
            \KwIn{  $S$: empty solution $S$, $K$: vehicles set \hspace{15mm} \\$\Pi$: custom sorted customers list }

@@

@@

For each delivery $d$ in $CL'$, the algorithm determines the best available driver $k$ and adds the corresponding loading and unloading task $(L^k_d, U^k_d)$ to a list of tasks. This is done using the $SimulateVisit$ procedure described earlier. Next, the algorithm constructs a $RCL$ from the list of tasks. If it is empty, the algorithm proceeds to the next iteration. Otherwise, it randomly selects a pair of tasks from the $RCL$, inserts them into $S$, and updates the remaining order and customer quantities. If an order $o$ is not fulfilled after visiting $d^j_{o}$, we add its next delivery node $d^{j+1}_{o}$ to $CL$.  Otherwise, if the customer has remaining orders, we select the first delivery node $d^{0}_{o'}$ of another randomly selected order $o'$ and add it to $CL$.

@@

\subsection{Local Search}

After the construction phase, the local search phase iteratively invokes five operators using a first-improvement strategy. The first operator (\textit{Swap Load}) exchanges loads between two deliveries of the same order. The \textit{Swap Driver} operator explores alternative driver assignments by swapping drivers assigned to two delivery nodes. The \textit{Relocate} operator removes a delivery node and places it in a different location. The \textit{Consolidate Load} operator assigns an order request initially split between two drivers to a single driver, if the driver's capacity allows. The \textit{Remove and Reschedule} operator selects an unscheduled customer, denoted $i$, within the current solution. It then identifies a set $C$ containing scheduled customers with the same time slot as $i$. All elements of $C$ are removed and rescheduled along with $i$. This local search process continues until no further improvements can be made or a timeout is reached.

@@

After the construction and local search steps, it is possible that some customers remain unscheduled in one solution while being accommodated in another. The goal of the path relinking procedure is to use each customer's scheduling information from existing solutions and combine them effectively to construct a solution that schedules all customers. This algorithm iterates through each element of a pool of solutions and attempts to insert each customer's schedule into a newly constructed solution.

\section{Computational results}
\label{comp_exp}
